The language of generalization: Probability, vagueness, and interaction

### 2018-03-26 analysis/case1-generics.Rmd
Finish scatterplot in appendix for parameter inference given different data sources

d.gen.endorse.bayes <- left_join(
  d.gen.endorse, 
  d.gen.endorse.catch %>% select(workerid, pass)
  ) %>%
  filter(pass == 1) %>%
  rowwise() %>%
  mutate(response = ifelse(response == "agree-key", 1, 0),
         sentence = gsub('&quotechar', '', sentence),
         sentence = gsub('lyme', 'Lyme', sentence)) %>%
  group_by(sentence) %>%
  summarize(k = sum(response), n = n()) %>%
  ungroup() %>%
  mutate(a = 1 + k,
         b = 1 + n - k,
         low  = qbeta(.025, a, b),
         high = qbeta(.975, a, b),
         MAP_h = (a-1)/(a+b-2),
         mean = a / (a + b)) %>% 
  ungroup()

@@

## Appendix C: Comparison of prior parameters inferred using only prior data and in joint-inference model
```{r generic-priors-parameters, eval = F}
m.gen.endorse.priors.posteriorPredictive <- data.frame()
n_samples <- 20000
lg = 0
# example.generics.prior.properties <- c(
#   "dont eat people", "have beautiful feathers",
#   "have wings", "are red",
#   "carry malaria", "lay eggs", 
#   "are female", "have spots"
#   )

@@

```{r}
left_join(
  m.gen.endorse.priors.parameters,
  m.gen.endorse.joint.prior.parameters %>%
    filter(type == "prior") %>%
    rowwise() %>%
    mutate(Parameter = ifelse(param == "mixture", "phi",
                              ifelse(category == "mean", "g", "d"))) %>%
    ungroup() %>% 
    select(property, Parameter, MAP, cred_upper, cred_lower) %>%
    rename(Property = property,
           MAP_joint = MAP,
           cred_upper_joint = cred_upper,
           cred_lower_joint = cred_lower)
) %>%
  mutate(Parameter = factor(Parameter,
                            levels = c("phi", "g", "d"),
                            labels = c(expression(paste("mixture parameter (", phi,")", sep = "")),
                                       expression(paste("stable component: mean (", gamma,")", sep = "")),
                                       expression(paste("stable component: concentration (", xi,")", sep = ""))))) %>%
  ggplot(., aes( x = MAP, xmin = cred_lower, xmax = cred_upper,
                 y = MAP_joint, ymin = cred_lower_joint, ymax = cred_upper_joint ))+
  geom_point()+
  geom_errorbar(alpha = 0.3)+geom_errorbarh(alpha = 0.3)+
  facet_wrap(~Parameter, scales = 'free', labeller = label_parsed)+
  xlab("Inferred values with only prevalence data")+
  ylab("Inferred values in joint-inference model")
```

>>>>

d.gen.endorse.bayes <- left_join(
  d.gen.endorse, 
  d.gen.endorse.catch %>% select(workerid, pass)
  ) %>%
  filter(pass == 1) %>%
  rowwise() %>%
  mutate(response = ifelse(response == "agree-key", 1, 0),
         sentence = gsub('&quotechar', '', sentence),
         sentence = gsub('lyme', 'Lyme', sentence)) %>%
  group_by(sentence, Property, Category) %>%
  summarize(k = sum(response), n = n()) %>%
  ungroup() %>%
  mutate(a = 1 + k,
         b = 1 + n - k,
         low  = qbeta(.025, a, b),
         high = qbeta(.975, a, b),
         MAP_h = (a-1)/(a+b-2),
         mean = a / (a + b)) %>% 
  ungroup()

@@

## Appendix C: Comparison of prior parameters inferred using only prior data and in joint-inference model

### Prevalence prior parameters

```{r generic-priors-parameters, eval = F}
m.gen.endorse.priors.posteriorPredictive <- data.frame()
n_samples <- 20000
lg = 0

@@

```{r}
load( file = '../analysis/generics-priors-posteriorParameters.Rdata')
m.gen.endorse.priors.parameters <- left_join(
  m.gen.endorse.priors.parameters,
  m.gen.endorse.joint.prior.parameters %>%
    filter(type == "prior") %>%
    rowwise() %>%
    mutate(Parameter = ifelse(param == "mixture", "phi",
                              ifelse(category == "mean", "g", "d"))) %>%
    ungroup() %>% 
    select(property, Parameter, MAP, cred_upper, cred_lower) %>%
    rename(Property = property,
           MAP_joint = MAP,
           cred_upper_joint = cred_upper,
           cred_lower_joint = cred_lower)
)

ggplot(m.gen.endorse.priors.parameters, aes( x = MAP, xmin = cred_lower, xmax = cred_upper,
                 y = MAP_joint, ymin = cred_lower_joint, ymax = cred_upper_joint ))+
  geom_point()+
  geom_errorbar(alpha = 0.3)+geom_errorbarh(alpha = 0.3)+
  facet_wrap(~Parameter, scales = 'free', labeller = label_parsed)+
  xlab("Inferred values with only prevalence data")+
  ylab("Inferred values in joint-inference model")
```

### Referent prevalence parameters

```{r generic-refprev-parameters, eval = F}
m.gen.endorse.refprev.posteriorPredictive <- data.frame()
n_samples <- 20000
lg = 0

m.gen.endorse.refprev.parameters <- data.frame()

for (p in levels(d.gen.endorse.priors$Property)) {
  categories <- d.gen.endorse.bayes %>% filter(Property == p) %>% distinct(Category)

  for (k in categories$Category) {

    priorData <- filter(d.gen.endorse.priors, Property == p, Category == k)$prevalence

    m.gen.refprev <- webppl(
      paste(genericEndorsementPriorModelHelpers, unstructuredPriorModel, sep = ""),
                                 data = priorData, data_var = "data",
                                 model_var = "model",
                                 chains = 2,
                                 cores = 2,
                                 inference_opts =
                                   list(method = "MCMC",
                                        samples = n_samples,
                                        burn = n_samples / 2,
                                        lag = lg,
                                        verbose = T))

  m.gen.refprev.summary <- m.gen.refprev %>%
    group_by(Parameter) %>%
    summarize(MAP = estimate_mode(value),
              cred_upper = hdi_upper(value),
              cred_lower = hdi_lower(value))

    m.gen.endorse.refprev.parameters <- bind_rows(
      m.gen.endorse.refprev.parameters,
      m.gen.refprev.summary %>%
        mutate(Property = p, Category = k)
      )
    print(k)
  }
  print(p)
}

 #save(m.gen.endorse.refprev.parameters, file = '../analysis/generics-refprev-posteriorParameters.Rdata')
```

```{r}
load(file = '../analysis/generics-refprev-posteriorParameters.Rdata')

m.gen.endorse.refprev.parameters <- left_join(
  m.gen.endorse.refprev.parameters,
  m.gen.endorse.joint.prior.parameters %>%
    filter(type == "targetPrevalence") %>%
    rowwise() %>%
    mutate(Parameter = ifelse(param == "mean", "g", "d")) %>%
    ungroup() %>% 
    select(property, category, Parameter, MAP, cred_upper, cred_lower) %>%
    rename(Property = property,
           Category = category,
           MAP_joint = MAP,
           cred_upper_joint = cred_upper,
           cred_lower_joint = cred_lower)
)

ggplot(m.gen.endorse.refprev.parameters, aes( x = MAP, xmin = cred_lower, xmax = cred_upper,
                 y = MAP_joint, ymin = cred_lower_joint, ymax = cred_upper_joint ))+
  geom_point()+
  geom_errorbar(alpha = 0.3)+geom_errorbarh(alpha = 0.3)+
  facet_wrap(~Parameter, scales = 'free', labeller = label_parsed)+
  xlab("Inferred parameters using only prevalence data")+
  ylab("Inferred parameters in joint-inference model")
```

```{r}
save(m.gen.endorse.priors.parameters, m.gen.endorse.refprev.parameters,
     file = "../paper/cached_results/case1_prevParamDistortion.RData")
```

### 2018-03-26 paper/appendix-bda.Rmd
Finish scatterplot in appendix for parameter inference given different data sources

```{r genericsModelDiagram, fig.cap="Pseudo-graphical model corresponding to the fully Bayesian data analysis of the endorsement model. The prevalence prior data $d^{pr}_{f}$ is assumed to be generated from the mixture model validated in Expt. 1a. The referent-prevalence $d^{pr}_{k,f}$ is generated from a latent prevalence $p_{fk}$. The fixed-threshold (control model) or uncertain-threshold model take the prevalence priors $P(p_{f})$, the referent-prevalence $p_{fk}$, and a single speaker optimality parameter $\\lambda$ as input. This overall structured is repeated (except $\\lambda$) for each of the unique properties $f$ and categories $k$ that correspond to the generic sentences in our stimulus set. Note that S corresponds to a probabilistic *function* and not a *random variable* that is standard in graphical model notation; S cannot be represented by a graphical model because it has recursion. "}
knitr::include_graphics("figs/generics_bayesnetmodel.pdf", dpi = 108)
```

@@

In addition to examining the posterior predictive distribution on endorsement judgments (presented in main text), we examined the marginal posteriors on parameters of the prevalence priors and referent-prevalence.
These marginal distributions are important to examine to confirm that they have not changed significantly from the parameters inferred from their private data sources in isolation.
For example, when modeling the referent-prevalence data in isolation, the model infers that roughly 65\% of robins lay eggs, as that is what participants on average produce in the prevalence elicitation task.^[
  Most participants report that 50\% of robins lay eggs, while a minority respond 100\%.
]
The question for this joint inference model is whether or not it still believes that roughly 50\% of robins lay eggs after being forced to jointly predict endorsements for "Robins lay eggs".
For example, a linear regression model based on referent-prevalence (of the kind presented in the main text) incorporated into this Bayesian analysis model will end up predicting high endorsement for "Robins lay eggs", but only does so by distorting the referent-prevalence data (inferring that 100\% of robins lay eggs); thus, the linear model sacrifices its goodness-of-fit to the referent-prevalence data in order to increase its goodness-of-fit to the endorsement data. 
Such a distortion would show manifest as a difference between the inferred parameters given only the referent-prevalence data and given the full joint model (all data sources modeled  simultaneously).

We found that the referent-prevalences and prevalence priors inferred under the joint model were almost indistinguishable from those inferred using only their data, suggesting that the modeling all three data sources together does not come at the cost of modeling any one of them accurately.

>>>>

```{r genericsModelDiagram, fig.cap="Quasi-graphical model corresponding to the fully Bayesian data analysis of the endorsement model. The prevalence prior data $d^{pr}_{f}$ is assumed to be generated from the mixture model validated in Expt. 1a. The referent-prevalence $d^{pr}_{k,f}$ is generated from a latent prevalence $p_{fk}$. The fixed-threshold (control model) or uncertain-threshold model take the prevalence priors $P(p_{f})$, the referent-prevalence $p_{fk}$, and a single speaker optimality parameter $\\lambda$ as input. This overall structured is repeated (except $\\lambda$) for each of the unique properties $f$ and categories $k$ that correspond to the generic sentences in our stimulus set. Note that S corresponds to a probabilistic *function* and not a *random variable* that is standard in graphical model notation; S cannot be represented by a graphical model because it has recursion. "}
knitr::include_graphics("figs/generics_bayesnetmodel.pdf", dpi = 108)
```

@@

In addition to examining the posterior predictive distribution on endorsement judgments (presented in main text), we examined the marginal posteriors on parameters of the prevalence priors and referent-prevalence.
These marginal distributions are important to confirm that they have not changed substantially from the parameters inferred from their primary data sources in isolation.
For example, when modeling the referent-prevalence data in isolation, the model infers that roughly 65\% of robins lay eggs, as that is what participants on average produce in the prevalence elicitation task.^[
  Most participants report that 50\% of robins lay eggs, while a minority respond 100\%.
]
If the joint inference model (which models all data sources---referent prevalence, prevalence prior, and generic endorsement---simultaneously) infers something substantially different about referent-prevalence than a model of referent prevalence in isolation, that could into question the inferences we as scientists derive from the joint inference model.
For example, incorporating a linear regression model (of the kinds presented as alterantive models in the main text) into this Bayesian joint-inference analysis model produces posterior predictions that match the generic endorsement data surprisingly well (e.g., that model predicts "Robins lay eggs" is true).
Such a model is only able to do this by distorting the referent-prevalence data, inferring that 100\% of robins lay eggs; thus, the linear model in this joint-inference analysis framework sacrifices its goodness-of-fit to the referent-prevalence data in order to increase its goodness-of-fit to the endorsement data. ^[
  This distortion effect is why we accout for measurement uncertainty in the linear models by bootstrapping the data that forms their predictors.
]
Such a distortion manifests as a difference between the inferred parameters given only the referent-prevalence data and given the full joint model (all data sources simultaneously).

```{r prevParamDistortion}
load("cached_results/case1_prevParamDistortion.RData") #m.gen.endorse.priors.parameters, m.gen.endorse.refprev.parameters

fig.distortion.prevpriors <- ggplot(
  m.gen.endorse.priors.parameters %>%
    mutate(Parameter = factor(Parameter,
                            levels = c("phi", "g", "d"),
                            labels = c(expression(paste("mixture parameter (", phi,")", sep = "")),
                                       expression(atop("stable component",
                                                  paste("mean (", gamma,")", sep = ""))),
                                       expression(atop("stable component", 
                                                  paste("concentration (", xi,")", sep = "")))))), 
  aes( x = MAP, xmin = cred_lower, xmax = cred_upper,
                 y = MAP_joint, ymin = cred_lower_joint, ymax = cred_upper_joint ))+
  geom_point()+
  geom_errorbar(alpha = 0.3)+geom_errorbarh(alpha = 0.3)+
  facet_wrap(~Parameter, scales = 'free', labeller = label_parsed)+
  xlab("Prevalence prior model parameters")+
  ylab("Joint-inference model parameters")

fig.distortion.refprev <-ggplot(
  m.gen.endorse.refprev.parameters %>%
    mutate(Parameter = factor(Parameter,
                            levels = c("g", "d"),
                            labels = c(expression(atop("referent prevalence", 
                                                       paste(" mean (", gamma,")", sep = ""))),
                                       expression(atop("referent prevalence",
                                                  paste("concentration (", xi,")", sep = "")))))), 
  aes( x = MAP, xmin = cred_lower, xmax = cred_upper,
                 y = MAP_joint, ymin = cred_lower_joint, ymax = cred_upper_joint ))+
  geom_point()+
  geom_errorbar(alpha = 0.3)+geom_errorbarh(alpha = 0.3)+
  facet_wrap(~Parameter, scales = 'free', labeller = label_parsed)+
  xlab("Referent prevalence model parameters")+
  ylab("Joint-inference model parameters")
```

```{r distortion, fig.width=10, fig.asp = 0.29, out.width = "100%", fig.align = "center", cache = F, fig.cap="No evidence for parameter distortion caused by a joint-inference model. X-axis denotes Maximum A-Posteriori (MAP) estimates inferred using only a single data source (A: Prevalence prior data. B: Referent prevalence data). Y-axis denotes MAP estimates inferred using the joint-inference model which models all data sources simultaneously. Distortions would manifest by systematic deviations from the y=x line."}
prow <- plot_grid( 
  fig.distortion.prevpriors + theme(plot.margin = unit(c(6,6,0,6), "pt")) #+ ggtitle("prevalence priors")
,
  fig.distortion.refprev + theme(axis.title.y = element_blank(),
                                 plot.margin = unit(c(6,0,0,0), "pt")),# + theme(legend.position="none"),
   align = 'h',
   labels = c("A", "B"),
   #hjust = -1,
   nrow = 1,
  rel_widths = c(1, 0.7),
  rel_heights = c(1, 0.7)
  )

prow
```

To investigate this distortion effect in the parameters, we compare the values inferred for the parameters governing the prevalence priors and referent prevalence variables before and after the joint-inference model sees the generic endorsement data.
Specifically, we infer the parameters for prevalence priors and referent prevalence by constructing single models of these tasks and compare the inferred values to those that result from the joint-inference model. 
The inferred parameters for these two models are shown in Figure\ \@ref(fig:distortion).
We found that the referent-prevalences and prevalence priors inferred under the joint model were almost indistinguishable from those inferred using only the referent-prevalence and prevalence prior data, respectively.
This results confirms that modeling all three data sources simultaneously does not distort some data sources (e.g., referent prevalence) in order to provide good fits for others (e.g., generic endorsement).

### 2018-03-26 paper/case1-generics.Rmd
Finish scatterplot in appendix for parameter inference given different data sources

<!-- Accurately modeling the prior data is important to generate precise quantitative predictions about endorsement. -->
In order to integrate the results of this prior measurement task into our model of endorsement, we must accurately model the prior elicitation data.
We approximate the prevalence distribution for each property (e.g., \textsc{lay eggs}) with a Mixture of Betas model, which assumes that the data generated for each kind comes from one of two underlying Beta distributions.^[
  The Beta distribution is chosen because the support of this distribution is numerical values between 0 - 1, exactly the form of the response data in the prior elicitation task.
] 
We specify one of these distributions *a priori* to represent kinds of animals who *do not have* a stable causal mechanism that could give rise to the property (e.g., \textsc{lions} and \textsc{lay eggs}), which results in prevalence or prevalence values close to or equal to 0.^[
  This assumption is similar in spirit that employed by *Hurdle Models* of epidemiological data, where the observed count of zeros is often substantially greater than one would expect from standard models, such as the Poisson [e.g., when modeling adverse reactions to vaccines; @hurdleModels]
]
This "null distribution" is potentially present for all features in exactly the same way (i.e., the lack of producing the feature).
The second distribution represents kinds of animals who *do have* such a mechanism, and the two parameters of this distribution are not specified *a priori* and are not the same for all properties, but are inferred on a property-wise basis from participants' responses.
The *Mixture of Betas* distribution has a third free parameter (for each property), the relative contribution of the "null distribution" (for example: we expect the "null distribution" to not contribute at all to properties like *being female*, for which almost all categories have at least some members with the property).
To ensure the *Mixture of Betas* model of the prior is not overly complex, we fit an additional model that represents only a single underlying distribution (*Single Beta*) for a comparison.
<!-- is flexible and can represent different shapes for different properties (e.g., the differences in prevalence judgments for \textsc{lay eggs} vs. \text{has wings}). -->
For more details about model implementation and inference, see Appendix C.

@@

We first examine the posterior predictive distribution on the prevalence prior and referent prevalence data to ensure that the joint-inference model does not distort these parameters at the service of predicting the endorsement data (e.g., such a distortion could manifest by the joint-inference model inferring that 100% of mosquitos carry malaria in order to predict that "Mosquitos carry malaria" is a good utterance).
This is an important step in model validation because it tells us that our model's predictions are derived from reasonable values of the parameters (e.g., that not all mosquitos carry malaria).
Importantly, the model recapitulates the prior elicitation data (e.g., the probability of carry malaria among various species) and the referent prevalence data (e.g., the prevalence of carrying malaria among mosquitos) as well as it did when the Bayesian data analysis model only included these data and their associated model (results section of Expt. 1b).
\ndg{point to a figure or give some stat to back up thi claim?}
This result confirms that the theoretically-interesting predictions of this model --- predictions of generic endorsement --- are based on intuitively meaningful model components (i.e., the shapes of the prevalence distributions). 
Finally, the inferred optimality parameter in the endorsement model (Eq. \ref{eq:S1}) is `r round(uncertain.threshold.s1opt[[1,"MAP"]], 2)` [`r round(uncertain.threshold.s1opt[[1,"cred_lower"]], 2)`, `r round(uncertain.threshold.s1opt[[1,"cred_upper"]], 2)`], a range consistent with the literature on similar models.

>>>>

<!-- Accurately modeling the prior data is important to generate precise quantitative predictions about endorsement. -->
In order to integrate the results of this prior measurement task into our model of endorsement, we must accurately model the prior elicitation data.
We approximate the prevalence distribution for each property (e.g., \textsc{lay eggs}) with a Mixture of Betas model, which assumes that the data generated for each kind comes from one of two underlying Beta distributions.^[
  The Beta distribution is chosen because the support of this distribution is numerical values between 0 - 1, exactly the form of the response data in the prior elicitation task.
] 
We specify one of these distributions *a priori* to represent kinds of animals who *do not have* a stable causal mechanism that could give rise to the property (e.g., \textsc{lions} and \textsc{lay eggs}), which results in prevalence or prevalence values close to or equal to 0.^[
  This assumption is similar in spirit to that employed by *Hurdle Models* of epidemiological data, where the observed count of zeros is often substantially greater than one would expect from standard models, such as the Poisson [e.g., when modeling adverse reactions to vaccines; @hurdleModels]
]
This "null distribution" is potentially present for all features in exactly the same way (i.e., the lack of producing the feature).
The second distribution represents kinds of animals who *do have* such a mechanism, and the two parameters of this distribution are not specified *a priori* and are not the same for all properties, but are inferred on a property-wise basis from participants' responses.
The *Mixture of Betas* distribution has a third free parameter (for each property), the relative contribution of the "null distribution" (for example: we expect the "null distribution" to not contribute at all to properties like *being female*, for which almost all categories have at least some members with the property).
To ensure the *Mixture of Betas* model of the prior is not overly complex, we fit an additional model that represents only a single underlying distribution (*Single Beta*) for a comparison.
<!-- is flexible and can represent different shapes for different properties (e.g., the differences in prevalence judgments for \textsc{lay eggs} vs. \text{has wings}). -->
For more details about model implementation and inference, see Appendix C.

@@

```{r distortionEffect}
load("cached_results/case1_prevParamDistortion.RData") #m.gen.endorse.priors.parameters, m.gen.endorse.refprev.parameters

r2.prevprior.distortion <- compute_r2(
  m.gen.endorse.priors.parameters, 
  "MAP", "MAP_joint", sigfigs = 4
)

r2.refprev.distortion <- compute_r2(
  m.gen.endorse.refprev.parameters, 
  "MAP", "MAP_joint", sigfigs = 4
)
```

We first examine the posterior predictive distribution on the prevalence prior and referent prevalence data to ensure that the joint-inference model does not distort these parameters at the service of predicting the endorsement data (e.g., such a distortion could manifest by the joint-inference model inferring that 100% of mosquitos carry malaria in order to predict that "Mosquitos carry malaria" is a good utterance).
This is an important step in model validation because it tells us that our model's predictions are derived from reasonable values of the parameters (e.g., that not all mosquitos carry malaria).
Importantly, the model recapitulates the prior elicitation data (e.g., the probability of carry malaria among various species) and the referent prevalence data (e.g., the prevalence of carrying malaria among mosquitos) as well as it did when the Bayesian data analysis model only included these data and their associated model ($r_{\text{prevalence prior parameters}}^2(`r length(m.gen.endorse.refprev.parameters[,"Parameter"])`) = `r r2.refprev.distortion`$; $r_{\text{referent prevalence parameters}}^2(`r length(m.gen.endorse.priors.parameters[,"Parameter"])`) = `r r2.prevprior.distortion`$; see Appendix D, Figure 14).
This result confirms that the theoretically-interesting predictions of this model --- predictions of generic endorsement --- are based on intuitively meaningful model components (i.e., the shapes of the prevalence distributions). 
Finally, the inferred optimality parameter in the endorsement model (Eq. \ref{eq:S1}) is `r round(uncertain.threshold.s1opt[[1,"MAP"]], 2)` [`r round(uncertain.threshold.s1opt[[1,"cred_lower"]], 2)`, `r round(uncertain.threshold.s1opt[[1,"cred_upper"]], 2)`], a range consistent with the literature on similar models.
